Continuous Deductive Reasoning over Streamed RDF Data

### 2015-07-14 content/contribution/continuous-datalog.tex

We consider the entailment of the WC-Datalog program $\Pi^{C,r}$ not
as a single stream of fact instances (as for C-Datalog), but in three
parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$
that is the entailment of the TA-Datalog program $\langle R,
F^0\rangle$, a stream $P$ of fact instances entailed by the fact
instances in $F^0$ and $S$, and a stream $N$ of fact instance
negations: $\langle R, F^0, S, r \rangle \models \langle E^0, P, N
\rangle$. $P$ and $N$ can be considered to be the continuous reasoning
equivalents of the streams produced by the \emph{IStream} and
\emph{DStream} operators respectively of the \emph{CQL} continuous
query language \citep{CQL}, which create streams based on the
additions to and deletions from a relation. Note that $P$ is an
in-order stream whose members are a subset of those in $E^S$ for an
equivalent C-Datalog programe, while $N$ may be an out-of-order
stream.
\end{definition}

@@

We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
\rangle$ to its TA-Datalog equivalent by taking the window $W^{S,r}_i$
over stream $S$ of range $r \in \mathbb{N}$ at instant $i \in
\mathbb{T}$; we refer to this as the {\em instantaneous state} of the
WC-Datalog program, written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i
\rangle$. As previously shown in \refdef{continuous datalog: ta
  datalog program}, this may then be further reduced to the Datalog
program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.

The instantaneous entailments of a $\Pi^{C,r}_i$ are the union of the set
of persistent entailments of ${\langle R, F^0 \rangle \vDash E^0}$
(which are guaranteed to be entailed by every state, as positive
Datalog is monotonic), and the set $E^T_{i}$ of transient entailments
whose justifications are at least partially contained in
$W^{S,r}_i$. We define $E^T_i$ as:
\[
E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
\]

>>>>

We consider the entailment of the WC-Datalog program $\Pi^{C,r}$ not
as a single stream of fact instances (as for C-Datalog), but in three
parts: a set of fact instances $E^0 = cons^T(\langle R, F^0\rangle)$
that is the entailment of the TA-Datalog program $\langle R,
F^0\rangle$; a stream $P$ of fact instances entailed by the rules $R$ and
the fact instances in $F^0$ and $S$, ordered by the instant at which they
became justified (i.e. their instants); and a stream $N$ of those same fact
instances, ordered by the instant at which they ceased to be justified:
$\langle R, F^0, S, r \rangle \models \langle E^0, P, N \rangle$. $P$
and $N$ can be considered to be the continuous reasoning equivalents of
the streams produced by the \emph{IStream} and\emph{DStream} operators
respectively of the \emph{CQL} continuous query language \citep{CQL},
which create streams based on the additions to and deletions from a
relation. Note that $P$ is an in-order stream whose members are a subset
of those in $E^S$ for an equivalent C-Datalog program, while $N$ may be
an out-of-order stream.

%%%FIXME
In order to

@@

We can reduce a WC-Datalog program $\Pi^{C,r} = \langle R, F^0, S, r
\rangle$ at an instant $i \in \mathbb{T}$ to its TA-Datalog equivalent by
taking the window $W^{S,r}_i$ over stream $S$ of range $r \in \mathbb{N}$;
we refer to this as the {\em instantaneous state} of the WC-Datalog program,
written as $\Pi^{C,r}_i = \langle R, F^0 \cup W^{S,r}_i \rangle$. As
previously shown in \refdef{continuous datalog: ta datalog program}, this
may then be further reduced to the Datalog program $\Pi = \langle R, strip(F^0 \cup W^{S,r}_i) \rangle$.

The instantaneous entailment of a $\Pi^{C,r}_i$ may be divided into two
semantically distinct portions: the \emph{persistent} portion is the
entailment of ${\langle R, F^0 \rangle \models E^0}$, as in a WC-Datalog
program, which is guaranteed to be part of the entailment of every state
due to the monotonicity of positive Datalog; the \emph{transitive} portion
is the set $E^T_{i}$ of entailed fact instances whose justifications are at
least partially contained in $W^{S,r}_i$. We define $E^T_i$ as:
\[
E^T_i = \{ x \mid \Pi^{C,r}_i \models x \land J^T \cup W^{S,r}_i \neq \varnothing \}
\]
